---
title: "lab_07"
author: "derek willis"
date: "10/21/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## You will need

* A Census API key

## Load libraries and establish settings

You'll need to load two packages for this: the tidyverse and tidycensus.

**Task** load these two packages

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the libraries you'll need
library(tidyverse)
library(tidycensus)
```

## Setup Census API

You'll need your Census API Key:

**Task** set your API Key if you haven't already

```{r echo=FALSE}
census_api_key("1bbc644ebc0424d5910b1ca8b809784d7379f2f2", install=TRUE)
```

## Load and view ACS variables

You'll also need to load the list of American Community Survey variables from the 2020 5-year estimates so you can identify the codes you need:

**Task** load the variables from the 2020 5-year ACS and save them to a dataframe

```{r}

acs <- load_variables(2020, "acs5")

```

## Answer questions

**Q1** What is the Maryland jurisdiction with the lowest median age, according to the ACS 2020 5-year estimates? You will need to find the variable for median age first.
**A1** Baltimore County is the jurisdiction with the lowest median age.

```{r}

md <- get_acs(geography = "county",
              variables = c(medinage = "B01002_001"),
              state = "MD",
              year = 2020)%>% arrange(estimate)

md

```

**Q2** Which Maryland counties have a median age of at least 50? You can use the dataframe you just made to answer this, but you must write code to display only those counties with a median age of at least 50.
**A2** Worcester County and Talbot County have median ages of at least 50.

```{r}

md <- get_acs(geography = "county",
              variables = c(medinage = "B01002_001"),
              state = "MD",
              year = 2020)%>% 
              filter(estimate >= "50") %>% 
              arrange(estimate)

md

```

**Q3** We're interested in knowing which Maryland counties have the highest percentage of 5-17 year-olds who speak Spanish at home and *also* speak English fluently. Start with the code you wrote for Q1 and alter it to show the total population of 5-17 year-olds by county. Then create a second dataframe to show the number of 5-17 year-olds by county living in homes where Spanish is spoken but who speak English "very well" according to the ACS. To do this, you will need to find the ACS variables that represent those subjects. HINT: the phrase "very well" shows up in multiple variables - you want the one that doesn't include other factors like citizenship. 

For each dataframe, rename the `estimate` column to a meaningful name you've given the ACS variable, remove the `variable` columns in each using select and then join them on GEOID and NAME. Then you can add a column to calculate the percentage of the total 5-17 population that speaks Spanish at home.

Which county has the highest percentage of 5-17 year-olds who speak Spanish at home and also speak English very well?

**A3** It appears that Prince George's County has the highest percentage of 5-17 year-olds who speak Spanish at home and also speak English very well, with almost 19%.

```{r}
md_young <- get_acs(geography = "county",
              variables = c(total = "B16004_002"),
              state = "MD",
              year = 2020)%>% arrange(estimate) %>% 
              mutate(hi = (estimate-moe), lo = (estimate+moe))

md_young

```
```{r}
md_span_eng <- get_acs(geography = "county",
              variables = c(total = "B16004_005"),
              state = "MD",
              year = 2020)%>% arrange(estimate) %>% 
              mutate(hi = (estimate-moe), lo = (estimate+moe))

md_span_eng

```
```{r}
md_young%>% 

  inner_join(md_span_eng, by=c("GEOID", "NAME")) %>% 
  
  select(GEOID, NAME, estimate.x, moe.x, estimate.y, moe.y) %>% 
  mutate(percent_lang = (estimate.y/estimate.x)*100) %>% 
  arrange(desc(percent_lang))
    
```
**Q4** Given the counties with the largest percentage of 5-17 year-olds that speak Spanish at home, which county stands out to you in the data as different from the others? How would you describe the top 5 counties? How would you describe Garrett County's numbers?

**A4** I saw that the first, second, and fifth counties are urban and relatively high in overall population, while the third and fourth are low in overall population. The margins of error are also quite high in some of the counties. Garrett county has the lowest percentage, and also has a small population overall.

**Q5** Pennsylvanians are currently requesting and submitting mail ballots for the November election. You'd like to know more about the people who live in the areas sending in mail ballots. Read in the `zip_totals.csv` file in the data folder containing the number of returned mail ballots per zip code. Then create a dataframe of Pennsylvania ZCTAs that include the following variables from the 2019 ACS: total population and median income. You will need to add `output = "wide"` to the get_acs function so that you get columns for those two variables' estimates and margins of error.

Join the ZCTA dataframe to the zip totals dataframe and calculate a per capita rate for mail ballots per 100 population in each ZCTA, and write code to answer the following questions?

* Which zip code *that has returned at least 100 mail ballots* has the highest per capita rate? 
* Where is that zip code and what does Census Reporter say about its demographics? 
* For zip codes with a per capita rate of at least 10 returned ballots per 100 people, how many have median incomes of at least $100,000? 
* How many have median incomes less than the state figure of $68,957?
* What's the most interesting one from this last result?

**A5**
A.19436, which is in Gwynedd, PA.

```{r}
pa_pop_income%>% 

  inner_join(pa, by=c("zip")) %>%
  select(zip, medinE, medinM, popE, popM, count) %>%
  mutate(per_100 = count/(popE/100)) %>% 
  arrange(desc(per_100))
  
    
```

B. 19436 is in Gwynedd, PA. The area has a population of 722 situated over 0.2 miles. It has a median age of 75.6, is 100% white and has a 57% female population. It's income per capita is $91,903, and median household income is $61,051.

C.I found none that had a median income of 100,000 or more.
```{r}
pa_100_cap%>% 

filter(medinE >= 100000)  
```

D. I found six zip codes that fit this description.

```{r}
pa_100_cap%>% 

filter(medinE >= 68957)  
```

E. The most interesting result in the last question was 15142, which is located in Allegheny County, PA. Not only does this locale have the highest number of returned ballots per 100 residents, it also has the lowest population and highest margin of error for median income.

```{r}

pa <- read_csv("data/pa_zip_totals.csv")

pa
```

```{r}

pa_pop_income <- get_acs(geography = "zcta",
              variables = c(medin = "B07011_001", 
                            pop = "B01003_001"),
              state = "PA",
              year = 2019, output = "wide") %>%
  rename(zip = GEOID)

pa_pop_income
```
```{r}
pa_100_cap <- pa_pop_income%>% 

  inner_join(pa, by=c("zip")) %>%
  select(zip, medinE, medinM, popE, popM, count) %>%
  mutate(per_100 = count/(popE/100)) %>% 
  filter(count >=100)%>%
  arrange(desc(medinE)) 
  
pa_100_cap   
```