---
title: "lab_02"
author: "derek willis"
date: "8/09/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and establish settings
```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse. If you have not installed the tidyverse already, remove the # from the next line and run it first.  
# install.packages('tidyverse')
library(tidyverse)
library(lubridate)
```

## Load Data

Let's keep working with the slice of WinRed contribution data from West Virginia.  

```{r}
wv_winred_contribs <- read_rds("data/lab_02.rds")
```

## Examine Data

## Answer questions

**Q1** Create a new column that divides the contribution amounts into the following options - less than $10, $10-$25, $26-$50, $51_$100, $101-$500, $501 and up - and calculates how many contributions are in each group. How would you describe the distribution of contributions in a sentence?
A1. Most contributions were made between $25 and $50, followed by contributions between $10 and $25, then by contributions over $500. 

```{r}

wv_winred_categories <- wv_winred_contribs %>%
  mutate(
    amount_category = case_when(
        amount < 10 ~ "Less than $10",
        amount >= 10 & amount <= 25 ~ "Between $10 and $25",
        amount > 25 & amount <= 50 ~ "Between $25 and $50",
        amount > 50 & amount <= 100 ~ "Between $50 and $100",
        amount > 100 & amount <= 500 ~ "Between $100 and $500",
        amount > 500 ~ "More than $500"
      )
  )
head(wv_winred_categories)

```
```{r}
wv_winred_categories %>% 
  group_by(amount_category) %>% 
  summarize(total_amount = sum(amount)) %>% 
  arrange(desc(total_amount))
```


Q2. Which West Virginia donors made the most contributions of less than $10? You'll need to use a combination of columns to get unique donors.  
A2. Betty Bender made the most contributions in that category.

```{r}

wv_winred_categories <- wv_winred_categories %>% filter(amount_category == "Less than $10")

head(wv_winred_categories)

```

Q3. What questions do the results of Q2 raise in your mind? How could you use the data to explore them? No code here, just your questions.
A3. My main question is about the fact that Betty Bender appears twice in the tibble, both contributions appear to be identical, so I would say that a good way to fix the discrepancy would be to run a code that deletes dublicate entries.

Q4. Which month saw the most contributions?
A4. April saw the most contributions.

```{r}
wv_winred_contribs %>%
  mutate(month = floor_date(date, "month")) %>%
  group_by(month) %>%
  summarise(total_amount = sum(amount)) %>%
  arrange(desc(total_amount))

```

Q5. How many contributions of $2,900 - the individual maximum donation - were there, and who did they go to? How many of the recipients are from West Virginia? Do some googling to find out.
A5. Six contributions were given of the amount of $2900, three of which were given to Carol for Congress. PALAZZO FOR CONGRESS is from Mississippi, ZINKE FOR CONGRESS is from Montana, and the other two campaigns are from West Virginia.

```{r}
wv_winred_contribs <- wv_winred_contribs %>% filter(aggregate_amount == "2900")

head(wv_winred_contribs)

```
